---
title: "Replication Report"
author: "Monica Chang"
date: "4/24/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(tidymodels)
library(readxl)
library(janitor)
library(rstanarm)
library(foreign)
library(stargazer)
library(arm)
library(haven)
library(lavaan)
library(modEvA)
library(dplyr)
library(xtable)
library(statxp)
library(gt)
```

## Overview of the original paper

## What I did for the replication

```{r clean data}
# I load the data from SPSS files from here:
# https://www.russellsage.org/datasets/economic-inequality-and-political-representation 

path = file.path("raw_data/DS1_v2_0.sav")
gilens = read_sav(path)

write_csv(gilens,"raw_data/DS1_v2_0.csv" )

# I clean the dataset to exclude questions about policy changes that would
# require a constitutional amendment or Supreme Court ruling, proposed changes 
# that were partially but not fully adopted, and questions that lacked income 
# breakdowns. 1,779 questions from 1981–2002 remain for the analyses.

clean_gilens <- gilens %>%
  clean_names() %>%
  
  # I exclude the 27 questions where there is no income breakdown.
  
  filter(!is.na(pred50_sw)) %>%
  
  # I exclude the 57 questions where the policy was partially adopted.
  
  filter(outcome != 2.5 & outcome != 3.5 & outcome != 4.5) 
  
  # None of the remaining questions require a Constitutional amendment or a 
  # new Supreme Court ruling to bring about the proposed policy change - so
  # I do not need to apply an additional filter.

```

```{r calculate and normalize predictors}
clean_gilens <- clean_gilens %>%
  
  # To use a logistic regression, I turn the outcome variable into a 
  # dichotomous variable.
  
  mutate(policy_change = if_else(outcome > 0, 1, 0)) %>%
  
  # I calculate the net interest group alignment score by using the formula 
  # described in the paper.
  
  mutate(intgrp_align = (log(intgrp_stfav + (0.5 * intgrp_swfav) + 1)) - 
                      (log(intgrp_stopp + (0.5 * intgrp_swopp) + 1))) %>%
  
  # I normalize the net interest group alignment score to be on a scale from 0 
  # to 1.
  
  mutate(intgrp_align_norm = range01(intgrp_align)) %>%
  
  # I use qlogis() to take the logit of the imputed percent of respondents at 
  # the 10th, 50th, and 90th income percentiles that favor the proposed policy 
  # change. I then use range01() to normalize my predictors to be on a scale 
  # from 0 to 1.
  
  mutate(pred10_norm = range01(qlogis(pred10_sw)),
         pred50_norm = range01(qlogis(pred50_sw)),
         pred90_norm = range01(qlogis(pred90_sw))) 
```

```{r add interest group alignment scores for mass-based & business interest groups}
clean_gilens <- clean_gilens %>%
  
  # I initialize 8 columns to track the number of mass-based and business
  # interest groups that are strongly in favor, somewhat in favor, somewhat
  # opposed, or strongly opposed to the policy change in each question.
  
  mutate(mass_intgrp_stfav = 0) %>%
  mutate(mass_intgrp_swfav = 0) %>%
  mutate(mass_intgrp_stopp = 0) %>%
  mutate(mass_intgrp_swopp = 0) %>%
  mutate(bus_intgrp_stfav = 0) %>%
  mutate(bus_intgrp_swfav = 0) %>%
  mutate(bus_intgrp_stopp = 0) %>%
  mutate(bus_intgrp_swopp = 0) %>%
  
  # I add 1 to the counter for each mass-based interest groups 
  # strongly in favor of the policy change in each question.
  
  mutate(mass_intgrp_stfav = ifelse(!is.na(aarp) & aarp == 2, mass_intgrp_stfav + 1, mass_intgrp_stfav)) %>%
  mutate(mass_intgrp_stfav = ifelse(!is.na(aflcio) & aflcio == 2, mass_intgrp_stfav + 1, mass_intgrp_stfav)) %>%
  mutate(mass_intgrp_stfav = ifelse(!is.na(american_federationof_state_countyand_municipal_employees) & american_federationof_state_countyand_municipal_employees == 2, mass_intgrp_stfav + 1, mass_intgrp_stfav)) %>%
  mutate(mass_intgrp_stfav = ifelse(!is.na(american_israel_public_affairs_committee) & american_israel_public_affairs_committee == 2, mass_intgrp_stfav + 1, mass_intgrp_stfav)) %>%
  mutate(mass_intgrp_stfav = ifelse(!is.na(american_legion) & american_legion == 2, mass_intgrp_stfav + 1, mass_intgrp_stfav)) %>%
  mutate(mass_intgrp_stfav = ifelse(!is.na(christian_coalition) & christian_coalition == 2, mass_intgrp_stfav + 1, mass_intgrp_stfav)) %>%
  mutate(mass_intgrp_stfav = ifelse(!is.na(international_brotherhoodof_teamsters) & international_brotherhoodof_teamsters == 2, mass_intgrp_stfav + 1, mass_intgrp_stfav)) %>%
  mutate(mass_intgrp_stfav = ifelse(!is.na(national_rifle_association) & national_rifle_association == 2, mass_intgrp_stfav + 1, mass_intgrp_stfav)) %>%
  mutate(mass_intgrp_stfav = ifelse(!is.na(national_rightto_life_committee) & national_rightto_life_committee == 2, mass_intgrp_stfav + 1, mass_intgrp_stfav)) %>%
  mutate(mass_intgrp_stfav = ifelse(!is.na(united_auto_workersunion) & united_auto_workersunion == 2, mass_intgrp_stfav + 1, mass_intgrp_stfav)) %>%
  mutate(mass_intgrp_stfav = ifelse(!is.na(veteransof_foreign_warsofthe_u_s) & veteransof_foreign_warsofthe_u_s == 2, mass_intgrp_stfav + 1, mass_intgrp_stfav)) %>%
  
  # I add 1 to the counter for each mass-based interest groups 
  # somewhat in favor of the policy change in each question.
  
  mutate(mass_intgrp_swfav = ifelse(!is.na(aarp) & aarp == 1, mass_intgrp_swfav + 1, mass_intgrp_swfav)) %>%
  mutate(mass_intgrp_swfav = ifelse(!is.na(aflcio) & aflcio == 1, mass_intgrp_swfav + 1, mass_intgrp_swfav)) %>%
  mutate(mass_intgrp_swfav = ifelse(!is.na(american_federationof_state_countyand_municipal_employees) & american_federationof_state_countyand_municipal_employees == 1, mass_intgrp_swfav + 1, mass_intgrp_swfav)) %>%
  mutate(mass_intgrp_swfav = ifelse(!is.na(american_israel_public_affairs_committee) & american_israel_public_affairs_committee == 1, mass_intgrp_swfav + 1, mass_intgrp_swfav)) %>%
  mutate(mass_intgrp_swfav = ifelse(!is.na(american_legion) & american_legion == 1, mass_intgrp_swfav + 1, mass_intgrp_swfav)) %>%
  mutate(mass_intgrp_swfav = ifelse(!is.na(christian_coalition) & christian_coalition == 1, mass_intgrp_swfav + 1, mass_intgrp_swfav)) %>%
  mutate(mass_intgrp_swfav = ifelse(!is.na(international_brotherhoodof_teamsters) & international_brotherhoodof_teamsters == 1, mass_intgrp_swfav + 1, mass_intgrp_swfav)) %>%
  mutate(mass_intgrp_swfav = ifelse(!is.na(national_rifle_association) & national_rifle_association == 1, mass_intgrp_swfav + 1, mass_intgrp_swfav)) %>%
  mutate(mass_intgrp_swfav = ifelse(!is.na(national_rightto_life_committee) & national_rightto_life_committee == 1, mass_intgrp_swfav + 1, mass_intgrp_swfav)) %>%
  mutate(mass_intgrp_swfav = ifelse(!is.na(united_auto_workersunion) & united_auto_workersunion == 1, mass_intgrp_swfav + 1, mass_intgrp_swfav)) %>%
  mutate(mass_intgrp_swfav = ifelse(!is.na(veteransof_foreign_warsofthe_u_s) & veteransof_foreign_warsofthe_u_s == 1, mass_intgrp_swfav + 1, mass_intgrp_swfav)) %>%
  
  # I add 1 to the counter for each mass-based interest groups 
  # strongly opposed to the policy change in each question.
  
  mutate(mass_intgrp_stopp = ifelse(!is.na(aarp) & aarp == -2, mass_intgrp_stopp + 1, mass_intgrp_stopp)) %>%
  mutate(mass_intgrp_stopp = ifelse(!is.na(aflcio) & aflcio == -2, mass_intgrp_stopp + 1, mass_intgrp_stopp)) %>%
  mutate(mass_intgrp_stopp = ifelse(!is.na(american_federationof_state_countyand_municipal_employees) & american_federationof_state_countyand_municipal_employees == -2, mass_intgrp_stopp + 1, mass_intgrp_stopp)) %>%
  mutate(mass_intgrp_stopp = ifelse(!is.na(american_israel_public_affairs_committee) & american_israel_public_affairs_committee == -2, mass_intgrp_stopp + 1, mass_intgrp_stopp)) %>%
  mutate(mass_intgrp_stopp = ifelse(!is.na(american_legion) & american_legion == -2, mass_intgrp_stopp + 1, mass_intgrp_stopp)) %>%
  mutate(mass_intgrp_stopp = ifelse(!is.na(christian_coalition) & christian_coalition == -2, mass_intgrp_stopp + 1, mass_intgrp_stopp)) %>%
  mutate(mass_intgrp_stopp = ifelse(!is.na(international_brotherhoodof_teamsters) & international_brotherhoodof_teamsters == -2, mass_intgrp_stopp + 1, mass_intgrp_stopp)) %>%
  mutate(mass_intgrp_stopp = ifelse(!is.na(national_rifle_association) & national_rifle_association == -2, mass_intgrp_stopp + 1, mass_intgrp_stopp)) %>%
  mutate(mass_intgrp_stopp = ifelse(!is.na(national_rightto_life_committee) & national_rightto_life_committee == -2, mass_intgrp_stopp + 1, mass_intgrp_stopp)) %>%
  mutate(mass_intgrp_stopp = ifelse(!is.na(united_auto_workersunion) & united_auto_workersunion == -2, mass_intgrp_stopp + 1, mass_intgrp_stopp)) %>%
  mutate(mass_intgrp_stopp = ifelse(!is.na(veteransof_foreign_warsofthe_u_s) & veteransof_foreign_warsofthe_u_s == -2, mass_intgrp_stopp + 1, mass_intgrp_stopp)) %>%
  
  # I add 1 to the counter for each mass-based interest groups 
  # somewhat opposed to the policy change in each question.
  
  mutate(mass_intgrp_swopp = ifelse(!is.na(aarp) & aarp == -1, mass_intgrp_swopp + 1, mass_intgrp_swopp)) %>%
  mutate(mass_intgrp_swopp = ifelse(!is.na(aflcio) & aflcio == -1, mass_intgrp_swopp + 1, mass_intgrp_swopp)) %>%
  mutate(mass_intgrp_swopp = ifelse(!is.na(american_federationof_state_countyand_municipal_employees) & american_federationof_state_countyand_municipal_employees == -1, mass_intgrp_swopp + 1, mass_intgrp_swopp)) %>%
  mutate(mass_intgrp_swopp = ifelse(!is.na(american_israel_public_affairs_committee) & american_israel_public_affairs_committee == -1, mass_intgrp_swopp + 1, mass_intgrp_swopp)) %>%
  mutate(mass_intgrp_swopp = ifelse(!is.na(american_legion) & american_legion == -1, mass_intgrp_swopp + 1, mass_intgrp_swopp)) %>%
  mutate(mass_intgrp_swopp = ifelse(!is.na(christian_coalition) & christian_coalition == -1, mass_intgrp_swopp + 1, mass_intgrp_swopp)) %>%
  mutate(mass_intgrp_swopp = ifelse(!is.na(international_brotherhoodof_teamsters) & international_brotherhoodof_teamsters == -1, mass_intgrp_swopp + 1, mass_intgrp_swopp)) %>%
  mutate(mass_intgrp_swopp = ifelse(!is.na(national_rifle_association) & national_rifle_association == -1, mass_intgrp_swopp + 1, mass_intgrp_swopp)) %>%
  mutate(mass_intgrp_swopp = ifelse(!is.na(national_rightto_life_committee) & national_rightto_life_committee == -1, mass_intgrp_swopp + 1, mass_intgrp_swopp)) %>%
  mutate(mass_intgrp_swopp = ifelse(!is.na(united_auto_workersunion) & united_auto_workersunion == -1, mass_intgrp_swopp + 1, mass_intgrp_swopp)) %>%
  mutate(mass_intgrp_swopp = ifelse(!is.na(veteransof_foreign_warsofthe_u_s) & veteransof_foreign_warsofthe_u_s == -1, mass_intgrp_swopp + 1, mass_intgrp_swopp)) %>%
  
  # I add 1 to the counter for each business interest groups 
  # strongly in favor of the policy change in each question.
  
  mutate(bus_intgrp_stfav = ifelse(!is.na(airlines) & airlines == 2, bus_intgrp_stfav + 1, bus_intgrp_stfav)) %>%
  mutate(bus_intgrp_stfav = ifelse(!is.na(american_bankers_association) & american_bankers_association == 2, bus_intgrp_stfav + 1, bus_intgrp_stfav)) %>%
  mutate(bus_intgrp_stfav = ifelse(!is.na(american_councilof_life_insurance) & american_councilof_life_insurance == 2, bus_intgrp_stfav + 1, bus_intgrp_stfav)) %>%
  mutate(bus_intgrp_stfav = ifelse(!is.na(american_farm_bureau_federation) & american_farm_bureau_federation == 2, bus_intgrp_stfav + 1, bus_intgrp_stfav)) %>%
  mutate(bus_intgrp_stfav = ifelse(!is.na(american_hospital_association) & american_hospital_association == 2, bus_intgrp_stfav + 1, bus_intgrp_stfav)) %>%
  mutate(bus_intgrp_stfav = ifelse(!is.na(american_medical_association) & american_medical_association == 2, bus_intgrp_stfav + 1, bus_intgrp_stfav)) %>%
  mutate(bus_intgrp_stfav = ifelse(!is.na(associationof_trial_lawyers) & associationof_trial_lawyers == 2, bus_intgrp_stfav + 1, bus_intgrp_stfav)) %>%
  mutate(bus_intgrp_stfav = ifelse(!is.na(automobilecompanies) & automobilecompanies == 2, bus_intgrp_stfav + 1, bus_intgrp_stfav)) %>%
  mutate(bus_intgrp_stfav = ifelse(!is.na(chamberof_commerce) & chamberof_commerce == 2, bus_intgrp_stfav + 1, bus_intgrp_stfav)) %>%
  mutate(bus_intgrp_stfav = ifelse(!is.na(computersoftwareandhardware) & computersoftwareandhardware == 2, bus_intgrp_stfav + 1, bus_intgrp_stfav)) %>%
  mutate(bus_intgrp_stfav = ifelse(!is.na(credit_union_national_association) & credit_union_national_association == 2, bus_intgrp_stfav + 1, bus_intgrp_stfav)) %>%
  mutate(bus_intgrp_stfav = ifelse(!is.na(defensecontractors) & defensecontractors == 2, bus_intgrp_stfav + 1, bus_intgrp_stfav)) %>%
  mutate(bus_intgrp_stfav = ifelse(!is.na(electriccompanies) & electriccompanies == 2, bus_intgrp_stfav + 1, bus_intgrp_stfav)) %>%
  mutate(bus_intgrp_stfav = ifelse(!is.na(health_insurance_association) & health_insurance_association == 2, bus_intgrp_stfav + 1, bus_intgrp_stfav)) %>%
  mutate(bus_intgrp_stfav = ifelse(!is.na(independent_insurance_agentsof_america) & independent_insurance_agentsof_america == 2, bus_intgrp_stfav + 1, bus_intgrp_stfav)) %>%
  mutate(bus_intgrp_stfav = ifelse(!is.na(motion_picture_associationof_america) & motion_picture_associationof_america == 2, bus_intgrp_stfav + 1, bus_intgrp_stfav)) %>%
  mutate(bus_intgrp_stfav = ifelse(!is.na(national_associationof_broadcasters) & national_associationof_broadcasters == 2, bus_intgrp_stfav + 1, bus_intgrp_stfav)) %>%
  mutate(bus_intgrp_stfav = ifelse(!is.na(national_associationof_home_builders) & national_associationof_home_builders == 2, bus_intgrp_stfav + 1, bus_intgrp_stfav)) %>%
  mutate(bus_intgrp_stfav = ifelse(!is.na(national_associationof_manufacturers) & national_associationof_manufacturers == 2, bus_intgrp_stfav + 1, bus_intgrp_stfav)) %>%
  mutate(bus_intgrp_stfav = ifelse(!is.na(national_associationof_realtors) & national_associationof_realtors == 2, bus_intgrp_stfav + 1, bus_intgrp_stfav)) %>%
  mutate(bus_intgrp_stfav = ifelse(!is.na(national_beer_wholesalers_association) & national_beer_wholesalers_association == 2, bus_intgrp_stfav + 1, bus_intgrp_stfav)) %>%
  mutate(bus_intgrp_stfav = ifelse(!is.na(national_federationof_independent_business) & national_federationof_independent_business == 2, bus_intgrp_stfav + 1, bus_intgrp_stfav)) %>%
  mutate(bus_intgrp_stfav = ifelse(!is.na(national_restaurant_association) & national_restaurant_association == 2, bus_intgrp_stfav + 1, bus_intgrp_stfav)) %>%
  mutate(bus_intgrp_stfav = ifelse(!is.na(oil_companies) & oil_companies == 2, bus_intgrp_stfav + 1, bus_intgrp_stfav)) %>%
  mutate(bus_intgrp_stfav = ifelse(!is.na(pharmaceutical_research_manufacturers) & pharmaceutical_research_manufacturers == 2, bus_intgrp_stfav + 1, bus_intgrp_stfav)) %>%
  mutate(bus_intgrp_stfav = ifelse(!is.na(recording_industry_association) & recording_industry_association == 2, bus_intgrp_stfav + 1, bus_intgrp_stfav)) %>%
  mutate(bus_intgrp_stfav = ifelse(!is.na(securitiesandinvestmentcompanies) & securitiesandinvestmentcompanies == 2, bus_intgrp_stfav + 1, bus_intgrp_stfav)) %>%
  mutate(bus_intgrp_stfav = ifelse(!is.na(telephonecompanies) & telephonecompanies == 2, bus_intgrp_stfav + 1, bus_intgrp_stfav)) %>%
  mutate(bus_intgrp_stfav = ifelse(!is.na(tobbacocompanies) & tobbacocompanies == 2, bus_intgrp_stfav + 1, bus_intgrp_stfav)) %>%
  
  # I add 1 to the counter for each business interest groups 
  # somewhat in favor of the policy change in each question.
  
  mutate(bus_intgrp_swfav = ifelse(!is.na(airlines) & airlines == 1, bus_intgrp_swfav + 1, bus_intgrp_swfav)) %>%
  mutate(bus_intgrp_swfav = ifelse(!is.na(american_bankers_association) & american_bankers_association == 1, bus_intgrp_swfav + 1, bus_intgrp_swfav)) %>%
  mutate(bus_intgrp_swfav = ifelse(!is.na(american_councilof_life_insurance) & american_councilof_life_insurance == 1, bus_intgrp_swfav + 1, bus_intgrp_swfav)) %>%
  mutate(bus_intgrp_swfav = ifelse(!is.na(american_farm_bureau_federation) & american_farm_bureau_federation == 1, bus_intgrp_swfav + 1, bus_intgrp_swfav)) %>%
  mutate(bus_intgrp_swfav = ifelse(!is.na(american_hospital_association) & american_hospital_association == 1, bus_intgrp_swfav + 1, bus_intgrp_swfav)) %>%
  mutate(bus_intgrp_swfav = ifelse(!is.na(american_medical_association) & american_medical_association == 1, bus_intgrp_swfav + 1, bus_intgrp_swfav)) %>%
  mutate(bus_intgrp_swfav = ifelse(!is.na(associationof_trial_lawyers) & associationof_trial_lawyers == 1, bus_intgrp_swfav + 1, bus_intgrp_swfav)) %>%
  mutate(bus_intgrp_swfav = ifelse(!is.na(automobilecompanies) & automobilecompanies == 1, bus_intgrp_swfav + 1, bus_intgrp_swfav)) %>%
  mutate(bus_intgrp_swfav = ifelse(!is.na(chamberof_commerce) & chamberof_commerce == 1, bus_intgrp_swfav + 1, bus_intgrp_swfav)) %>%
  mutate(bus_intgrp_swfav = ifelse(!is.na(computersoftwareandhardware) & computersoftwareandhardware == 1, bus_intgrp_swfav + 1, bus_intgrp_swfav)) %>%
  mutate(bus_intgrp_swfav = ifelse(!is.na(credit_union_national_association) & credit_union_national_association == 1, bus_intgrp_swfav + 1, bus_intgrp_swfav)) %>%
  mutate(bus_intgrp_swfav = ifelse(!is.na(defensecontractors) & defensecontractors == 1, bus_intgrp_swfav + 1, bus_intgrp_swfav)) %>%
  mutate(bus_intgrp_swfav = ifelse(!is.na(electriccompanies) & electriccompanies == 1, bus_intgrp_swfav + 1, bus_intgrp_swfav)) %>%
  mutate(bus_intgrp_swfav = ifelse(!is.na(health_insurance_association) & health_insurance_association == 1, bus_intgrp_swfav + 1, bus_intgrp_swfav)) %>%
  mutate(bus_intgrp_swfav = ifelse(!is.na(independent_insurance_agentsof_america) & independent_insurance_agentsof_america == 1, bus_intgrp_swfav + 1, bus_intgrp_swfav)) %>%
  mutate(bus_intgrp_swfav = ifelse(!is.na(motion_picture_associationof_america) & motion_picture_associationof_america == 1, bus_intgrp_swfav + 1, bus_intgrp_swfav)) %>%
  mutate(bus_intgrp_swfav = ifelse(!is.na(national_associationof_broadcasters) & national_associationof_broadcasters == 1, bus_intgrp_swfav + 1, bus_intgrp_swfav)) %>%
  mutate(bus_intgrp_swfav = ifelse(!is.na(national_associationof_home_builders) & national_associationof_home_builders == 1, bus_intgrp_swfav + 1, bus_intgrp_swfav)) %>%
  mutate(bus_intgrp_swfav = ifelse(!is.na(national_associationof_manufacturers) & national_associationof_manufacturers == 1, bus_intgrp_swfav + 1, bus_intgrp_swfav)) %>%
  mutate(bus_intgrp_swfav = ifelse(!is.na(national_associationof_realtors) & national_associationof_realtors == 1, bus_intgrp_swfav + 1, bus_intgrp_swfav)) %>%
  mutate(bus_intgrp_swfav = ifelse(!is.na(national_beer_wholesalers_association) & national_beer_wholesalers_association == 1, bus_intgrp_swfav + 1, bus_intgrp_swfav)) %>%
  mutate(bus_intgrp_swfav = ifelse(!is.na(national_federationof_independent_business) & national_federationof_independent_business == 1, bus_intgrp_swfav + 1, bus_intgrp_swfav)) %>%
  mutate(bus_intgrp_swfav = ifelse(!is.na(national_restaurant_association) & national_restaurant_association == 1, bus_intgrp_swfav + 1, bus_intgrp_swfav)) %>%
  mutate(bus_intgrp_swfav = ifelse(!is.na(oil_companies) & oil_companies == 1, bus_intgrp_swfav + 1, bus_intgrp_swfav)) %>%
  mutate(bus_intgrp_swfav = ifelse(!is.na(pharmaceutical_research_manufacturers) & pharmaceutical_research_manufacturers == 1, bus_intgrp_swfav + 1, bus_intgrp_swfav)) %>%
  mutate(bus_intgrp_swfav = ifelse(!is.na(recording_industry_association) & recording_industry_association == 1, bus_intgrp_swfav + 1, bus_intgrp_swfav)) %>%
  mutate(bus_intgrp_swfav = ifelse(!is.na(securitiesandinvestmentcompanies) & securitiesandinvestmentcompanies == 1, bus_intgrp_swfav + 1, bus_intgrp_swfav)) %>%
  mutate(bus_intgrp_swfav = ifelse(!is.na(telephonecompanies) & telephonecompanies == 1, bus_intgrp_swfav + 1, bus_intgrp_swfav)) %>%
  mutate(bus_intgrp_swfav = ifelse(!is.na(tobbacocompanies) & tobbacocompanies == 1, bus_intgrp_swfav + 1, bus_intgrp_swfav)) %>%
  
  # I add 1 to the counter for each business interest groups 
  # strongly opposed to the policy change in each question.
  
  mutate(bus_intgrp_stopp = ifelse(!is.na(airlines) & airlines == -2, bus_intgrp_stopp + 1, bus_intgrp_stopp)) %>%
  mutate(bus_intgrp_stopp = ifelse(!is.na(american_bankers_association) & american_bankers_association == -2, bus_intgrp_stopp + 1, bus_intgrp_stopp)) %>%
  mutate(bus_intgrp_stopp = ifelse(!is.na(american_councilof_life_insurance) & american_councilof_life_insurance == -2, bus_intgrp_stopp + 1, bus_intgrp_stopp)) %>%
  mutate(bus_intgrp_stopp = ifelse(!is.na(american_farm_bureau_federation) & american_farm_bureau_federation == -2, bus_intgrp_stopp + 1, bus_intgrp_stopp)) %>%
  mutate(bus_intgrp_stopp = ifelse(!is.na(american_hospital_association) & american_hospital_association == -2, bus_intgrp_stopp + 1, bus_intgrp_stopp)) %>%
  mutate(bus_intgrp_stopp = ifelse(!is.na(american_medical_association) & american_medical_association == -2, bus_intgrp_stopp + 1, bus_intgrp_stopp)) %>%
  mutate(bus_intgrp_stopp = ifelse(!is.na(associationof_trial_lawyers) & associationof_trial_lawyers == -2, bus_intgrp_stopp + 1, bus_intgrp_stopp)) %>%
  mutate(bus_intgrp_stopp = ifelse(!is.na(automobilecompanies) & automobilecompanies == -2, bus_intgrp_stopp + 1, bus_intgrp_stopp)) %>%
  mutate(bus_intgrp_stopp = ifelse(!is.na(chamberof_commerce) & chamberof_commerce == -2, bus_intgrp_stopp + 1, bus_intgrp_stopp)) %>%
  mutate(bus_intgrp_stopp = ifelse(!is.na(computersoftwareandhardware) & computersoftwareandhardware == -2, bus_intgrp_stopp + 1, bus_intgrp_stopp)) %>%
  mutate(bus_intgrp_stopp = ifelse(!is.na(credit_union_national_association) & credit_union_national_association == -2, bus_intgrp_stopp + 1, bus_intgrp_stopp)) %>%
  mutate(bus_intgrp_stopp = ifelse(!is.na(defensecontractors) & defensecontractors == -2, bus_intgrp_stopp + 1, bus_intgrp_stopp)) %>%
  mutate(bus_intgrp_stopp = ifelse(!is.na(electriccompanies) & electriccompanies == -2, bus_intgrp_stopp + 1, bus_intgrp_stopp)) %>%
  mutate(bus_intgrp_stopp = ifelse(!is.na(health_insurance_association) & health_insurance_association == -2, bus_intgrp_stopp + 1, bus_intgrp_stopp)) %>%
  mutate(bus_intgrp_stopp = ifelse(!is.na(independent_insurance_agentsof_america) & independent_insurance_agentsof_america == -2, bus_intgrp_stopp + 1, bus_intgrp_stopp)) %>%
  mutate(bus_intgrp_stopp = ifelse(!is.na(motion_picture_associationof_america) & motion_picture_associationof_america == -2, bus_intgrp_stopp + 1, bus_intgrp_stopp)) %>%
  mutate(bus_intgrp_stopp = ifelse(!is.na(national_associationof_broadcasters) & national_associationof_broadcasters == -2, bus_intgrp_stopp + 1, bus_intgrp_stopp)) %>%
  mutate(bus_intgrp_stopp = ifelse(!is.na(national_associationof_home_builders) & national_associationof_home_builders == -2, bus_intgrp_stopp + 1, bus_intgrp_stopp)) %>%
  mutate(bus_intgrp_stopp = ifelse(!is.na(national_associationof_manufacturers) & national_associationof_manufacturers == -2, bus_intgrp_stopp + 1, bus_intgrp_stopp)) %>%
  mutate(bus_intgrp_stopp = ifelse(!is.na(national_associationof_realtors) & national_associationof_realtors == -2, bus_intgrp_stopp + 1, bus_intgrp_stopp)) %>%
  mutate(bus_intgrp_stopp = ifelse(!is.na(national_beer_wholesalers_association) & national_beer_wholesalers_association == -2, bus_intgrp_stopp + 1, bus_intgrp_stopp)) %>%
  mutate(bus_intgrp_stopp = ifelse(!is.na(national_federationof_independent_business) & national_federationof_independent_business == -2, bus_intgrp_stopp + 1, bus_intgrp_stopp)) %>%
  mutate(bus_intgrp_stopp = ifelse(!is.na(national_restaurant_association) & national_restaurant_association == -2, bus_intgrp_stopp + 1, bus_intgrp_stopp)) %>%
  mutate(bus_intgrp_stopp = ifelse(!is.na(oil_companies) & oil_companies == -2, bus_intgrp_stopp + 1, bus_intgrp_stopp)) %>%
  mutate(bus_intgrp_stopp = ifelse(!is.na(pharmaceutical_research_manufacturers) & pharmaceutical_research_manufacturers == -2, bus_intgrp_stopp + 1, bus_intgrp_stopp)) %>%
  mutate(bus_intgrp_stopp = ifelse(!is.na(recording_industry_association) & recording_industry_association == -2, bus_intgrp_stopp + 1, bus_intgrp_stopp)) %>%
  mutate(bus_intgrp_stopp = ifelse(!is.na(securitiesandinvestmentcompanies) & securitiesandinvestmentcompanies == -2, bus_intgrp_stopp + 1, bus_intgrp_stopp)) %>%
  mutate(bus_intgrp_stopp = ifelse(!is.na(telephonecompanies) & telephonecompanies == -2, bus_intgrp_stopp + 1, bus_intgrp_stopp)) %>%
  mutate(bus_intgrp_stopp = ifelse(!is.na(tobbacocompanies) & tobbacocompanies == -2, bus_intgrp_stopp + 1, bus_intgrp_stopp)) %>%
  
  # I add 1 to the counter for each business interest groups 
  # somewhat opposed to the policy change in each question.
  
  mutate(bus_intgrp_swopp = ifelse(!is.na(airlines) & airlines == -1, bus_intgrp_swopp + 1, bus_intgrp_swopp)) %>%
  mutate(bus_intgrp_swopp = ifelse(!is.na(american_bankers_association) & american_bankers_association == -1, bus_intgrp_swopp + 1, bus_intgrp_swopp)) %>%
  mutate(bus_intgrp_swopp = ifelse(!is.na(american_councilof_life_insurance) & american_councilof_life_insurance == -1, bus_intgrp_swopp + 1, bus_intgrp_swopp)) %>%
  mutate(bus_intgrp_swopp = ifelse(!is.na(american_farm_bureau_federation) & american_farm_bureau_federation == -1, bus_intgrp_swopp + 1, bus_intgrp_swopp)) %>%
  mutate(bus_intgrp_swopp = ifelse(!is.na(american_hospital_association) & american_hospital_association == -1, bus_intgrp_swopp + 1, bus_intgrp_swopp)) %>%
  mutate(bus_intgrp_swopp = ifelse(!is.na(american_medical_association) & american_medical_association == -1, bus_intgrp_swopp + 1, bus_intgrp_swopp)) %>%
  mutate(bus_intgrp_swopp = ifelse(!is.na(associationof_trial_lawyers) & associationof_trial_lawyers == -1, bus_intgrp_swopp + 1, bus_intgrp_swopp)) %>%
  mutate(bus_intgrp_swopp = ifelse(!is.na(automobilecompanies) & automobilecompanies == -1, bus_intgrp_swopp + 1, bus_intgrp_swopp)) %>%
  mutate(bus_intgrp_swopp = ifelse(!is.na(chamberof_commerce) & chamberof_commerce == -1, bus_intgrp_swopp + 1, bus_intgrp_swopp)) %>%
  mutate(bus_intgrp_swopp = ifelse(!is.na(computersoftwareandhardware) & computersoftwareandhardware == -1, bus_intgrp_swopp + 1, bus_intgrp_swopp)) %>%
  mutate(bus_intgrp_swopp = ifelse(!is.na(credit_union_national_association) & credit_union_national_association == -1, bus_intgrp_swopp + 1, bus_intgrp_swopp)) %>%
  mutate(bus_intgrp_swopp = ifelse(!is.na(defensecontractors) & defensecontractors == -1, bus_intgrp_swopp + 1, bus_intgrp_swopp)) %>%
  mutate(bus_intgrp_swopp = ifelse(!is.na(electriccompanies) & electriccompanies == -1, bus_intgrp_swopp + 1, bus_intgrp_swopp)) %>%
  mutate(bus_intgrp_swopp = ifelse(!is.na(health_insurance_association) & health_insurance_association == -1, bus_intgrp_swopp + 1, bus_intgrp_swopp)) %>%
  mutate(bus_intgrp_swopp = ifelse(!is.na(independent_insurance_agentsof_america) & independent_insurance_agentsof_america == -1, bus_intgrp_swopp + 1, bus_intgrp_swopp)) %>%
  mutate(bus_intgrp_swopp = ifelse(!is.na(motion_picture_associationof_america) & motion_picture_associationof_america == -1, bus_intgrp_swopp + 1, bus_intgrp_swopp)) %>%
  mutate(bus_intgrp_swopp = ifelse(!is.na(national_associationof_broadcasters) & national_associationof_broadcasters == -1, bus_intgrp_swopp + 1, bus_intgrp_swopp)) %>%
  mutate(bus_intgrp_swopp = ifelse(!is.na(national_associationof_home_builders) & national_associationof_home_builders == -1, bus_intgrp_swopp + 1, bus_intgrp_swopp)) %>%
  mutate(bus_intgrp_swopp = ifelse(!is.na(national_associationof_manufacturers) & national_associationof_manufacturers == -1, bus_intgrp_swopp + 1, bus_intgrp_swopp)) %>%
  mutate(bus_intgrp_swopp = ifelse(!is.na(national_associationof_realtors) & national_associationof_realtors == -1, bus_intgrp_swopp + 1, bus_intgrp_swopp)) %>%
  mutate(bus_intgrp_swopp = ifelse(!is.na(national_beer_wholesalers_association) & national_beer_wholesalers_association == -1, bus_intgrp_swopp + 1, bus_intgrp_swopp)) %>%
  mutate(bus_intgrp_swopp = ifelse(!is.na(national_federationof_independent_business) & national_federationof_independent_business == -1, bus_intgrp_swopp + 1, bus_intgrp_swopp)) %>%
  mutate(bus_intgrp_swopp = ifelse(!is.na(national_restaurant_association) & national_restaurant_association == -1, bus_intgrp_swopp + 1, bus_intgrp_swopp)) %>%
  mutate(bus_intgrp_swopp = ifelse(!is.na(oil_companies) & oil_companies == -1, bus_intgrp_swopp + 1, bus_intgrp_swopp)) %>%
  mutate(bus_intgrp_swopp = ifelse(!is.na(pharmaceutical_research_manufacturers) & pharmaceutical_research_manufacturers == -1, bus_intgrp_swopp + 1, bus_intgrp_swopp)) %>%
  mutate(bus_intgrp_swopp = ifelse(!is.na(recording_industry_association) & recording_industry_association == -1, bus_intgrp_swopp + 1, bus_intgrp_swopp)) %>%
  mutate(bus_intgrp_swopp = ifelse(!is.na(securitiesandinvestmentcompanies) & securitiesandinvestmentcompanies == -1, bus_intgrp_swopp + 1, bus_intgrp_swopp)) %>%
  mutate(bus_intgrp_swopp = ifelse(!is.na(telephonecompanies) & telephonecompanies == -1, bus_intgrp_swopp + 1, bus_intgrp_swopp)) %>%
  mutate(bus_intgrp_swopp = ifelse(!is.na(tobbacocompanies) & tobbacocompanies == -1, bus_intgrp_swopp + 1, bus_intgrp_swopp)) %>%
  
  # I calculate the net interest group alignment scores for mass-based interest
  # groups and business interest groups by using the formula described in the 
  # paper.
  
  mutate(mass_intgrp_align = (log(mass_intgrp_stfav + (0.5 * mass_intgrp_swfav) + 1)) - 
                      (log(mass_intgrp_stopp + (0.5 * mass_intgrp_swopp) + 1)),
         bus_intgrp_align = (log(bus_intgrp_stfav + (0.5 * bus_intgrp_swfav) + 1)) - 
                      (log(bus_intgrp_stopp + (0.5 * bus_intgrp_swopp) + 1))) %>%
  
  # I normalize the net interest group alignment scores for mass-based interest
  # groups and business interest groups to be on a scale from 0 to 1.
  
  mutate(mass_intgrp_align_norm = range01(mass_intgrp_align),
         bus_intgrp_align_norm = range01(bus_intgrp_align)) 
```

```{r generate table 1}
# I initialize the appropriate values for sets of actors that would 
# have independent influence for each of the theories of U.S. politics.

table_1 <- tibble(theories = c("Majoritarian Electoral Democracy", 
                               "Dominance by Economic Elites", 
                               "Majoritarian Pluralism", 
                               "Biased Pluralism"),
                  pred50_norm = c("Y", "y", "y", "n"),
                  pred90_norm = c("n", "Y", "n", "n"),
                  intgrp_align_norm = c("n", "y", "Y", "y"),
                  mass_intgrp_align_norm = c("n", "n", "Y", "y"),
                  bus_intgrp_align_norm = c("n", "y", "Y", "Y"))

# I format the table using gt() functions.

table_1 %>%
  gt(rowname_col = "theories") %>%
  tab_header(
    title = "Table 1",
    subtitle = glue::glue("Theoretical predictions concerning the independent influence of sets of actors upon policy outcomes")
  ) %>%
  tab_spanner(
    label = "Sets of Actors",
    columns = c("pred50_norm", "pred90_norm", "intgrp_align_norm", "mass_intgrp_align_norm", "bus_intgrp_align_norm")
  ) %>%
  cols_label(pred50_norm = "Average Citizens",
             pred90_norm = "Economic Elites",
             intgrp_align_norm = "All Interest Groups",
             mass_intgrp_align_norm = "Mass Interest Groups",
             bus_intgrp_align_norm = "Business Interest Groups") %>%
  tab_source_note(
    source_note = "n = little or no independent influence"
  ) %>%
  tab_stubhead(label = "Theory (ideal type)") %>%
  tab_source_note(
    source_note = "y = some independent influence"
  ) %>%
  tab_source_note(
    source_note = "Y = substantial independent influence"
  )
```

```{r generate table 2}
# I select the five predictors relevant for Table 2.

table_2 <- clean_gilens %>%
  dplyr::select(pred50_norm, pred90_norm, intgrp_align_norm, 
                mass_intgrp_align_norm, bus_intgrp_align_norm)

# I use the corstars() function to compute a correlation matrix with stars to
# indicate significant correlations.

table_2_correlation <- corstars(table_2) %>%
  mutate(predictors = c("Average citizens’ preferences", "Economic elites’ preferences", 
                        "All interest groups", "Mass public interest groups", 
                        "Business interest groups"))

# I format the table using gt() functions.

table_2_correlation %>%
  gt(rowname_col = "predictors") %>%
  tab_header(
    title = "Table 2",
    subtitle = glue::glue("Correlations among independent variables")
  ) %>%
  cols_label(pred50_norm = "Average citizens’ preferences",
             pred90_norm = "Economic elites’ preferences",
             intgrp_align_norm = "All interest groups",
             mass_intgrp_align_norm = "Mass public interest groups") %>%
  tab_source_note(
    source_note = "***p<.001; n = 1779."
  ) %>%
  tab_source_note(
    source_note = "Note: Entries are correlation coefficients based on OLS regression coefficients - not SEM modeling."
  )
```

```{r generate table 3}
# I run a model where the preferences of the median voter is the sole predictor 
# of policy change.

model_1 <- lm(policy_change ~ pred50_norm, data = clean_gilens)
summary(model_1)

# I run a model where the preferences of the 90th percentile voter is the sole 
# predictor of policy change.

model_2 <- lm(policy_change ~ pred90_norm, data = clean_gilens)
summary(model_2)

# I run a model where the interest group alignment is the sole predictor of 
# policy change.

model_3 <- lm(policy_change ~ intgrp_align_norm, data = clean_gilens)
summary(model_3)

# I run a multivariate model where the preferences of the median voter, the 
# preferences of the 90th percentile voter, and the interest group alignment
# are predictors of policy change.

model_4 <- lm(policy_change ~ pred50_norm + pred90_norm + intgrp_align_norm, data = clean_gilens)
summary(model_4)

# I use stargazer() to format a table with 3 regressions side-by-side.

stargazer(model_1, model_2, model_3, model_4, 
          type = "html", 
          title = "Table 3. Policy outcomes and the policy preferences of average citizens, economic elites, and interest groups",
          align = TRUE,
          dep.var.labels = c("Policy change"),
          column.labels = c("Model 1", "Model 2", "Model 3", "Model 4"),
          covariate.labels = c("Preferences of average citizens", 
                               "Preferences of economic elites", 
                               "Alignment of interest groups"),
          omit.stat=c("LL","ser","f", "adj.rsq", "n"), 
          no.space=TRUE,
          notes = c("*p<0.1; **p<0.05; ***p<0.01", "All predictors are scaled to range from 0 to 1. The dependent variable is the policy outcome, coded 1 if the proposed policy change took place within four years of the survey date and 0 if it did not. Predictors are the logits of the imputed percent of respondents at the fiftieth (“average citizens”) or ninetieth (“economic elites”) income percentile that favor the proposed policy change, and the Net Interest- Group Alignment Index described in the text. Standard errors are asymptotically distribution-free, and all analyses reflect estimated measurement error in the predictors, as described in Appendix 2. The standardized coefficients for model 4 in this table are .01, .21, and .16 for average citizens, economic elites, and interest groups, respectively. N = 1,779."),
          notes.align = "l",
          notes.append = FALSE)
```


```{r figure 1}

clean_gilens %>%
  
  # I multiply pred50_sw to obtain a number that represents the percentage 
  # of median voters in favor of a proposed policy change.
  
  mutate(pred50_sw_percent = 100*pred50_sw) %>%
  ggplot(aes(pred50_sw_percent)) +
  
    # I use a histogram to display the underlying preference distribution.
  
    geom_histogram(aes(y = after_stat(count/sum(count))),
                   alpha = 0.5, 
                   bins = 10, 
                   position = "identity") +
  
    # I use the coefficients from my Model 4 output and the neutral points of 
    # the other two actors - 0.5 for economic elites and 0 for interest group
    # alignment - to construct a line graph that shows how the predicted 
    # probability of policy change alters when moving from one point to another 
    # on the distribution of policy dispositions.
  
    geom_line(aes(y = -0.20999 + -0.86716*pred50_norm + 1.60551*0.5 + 0.48878*0)) +
    xlim(0, 100) +
    
    # I customize the Y scales to have dual axes.
  
    scale_y_continuous(
      
      # I specify the name and limits of the lefthand axis.
      
      name = "Predicted probability of adoption",
      limits = c(0, 0.7),
      
      # I specify the name and limits of the second axis by transforming
      # the first axis.
      
      sec.axis = sec_axis(trans = ~.*0.5714, 
                          name = "Percent of cases (grey columns)")
    ) +
    labs(title = "Average Citizens' Preferences",
         x = "Percent favoring proposed policy change") +
    theme_classic()

clean_gilens %>%
  
  # I multiply pred90_sw to obtain a number that represents the percentage 
  # of median voters in favor of a proposed policy change.
  
  mutate(pred90_sw_percent = 100*pred90_sw) %>%
  ggplot(aes(pred90_sw_percent)) +
  
    # I use a histogram to display the underlying preference distribution.
  
    geom_histogram(aes(y = after_stat(count/sum(count))),
                   alpha = 0.5, 
                   bins = 10, 
                   position = "identity") +
  
    # I use the coefficients from my Model 4 output and the neutral points of 
    # the other two actors - 0.5 for average citizens and 0 for interest group
    # alignment - to construct a line graph that shows how the predicted 
    # probability of policy change alters when moving from one point to another 
    # on the distribution of policy dispositions.
  
    geom_line(aes(y = -0.20999 + -0.86716*0.5 + 1.60551*pred90_norm + 0.48878*0)) +
    xlim(0, 100) +
    
    # I customize the Y scales to have dual axes.
  
    scale_y_continuous(
      
      # I specify the name and limits of the lefthand axis.
      
      name = "Predicted probability of adoption",
      limits = c(0, 0.7),
      
      # I specify the name and limits of the second axis by transforming
      # the first axis.
      
      sec.axis = sec_axis(trans = ~.*0.5714, 
                          name = "Percent of cases (grey columns)")
    ) +
    labs(title = "Economic Elites' Preferences",
         x = "Percent favoring proposed policy change") +
    theme_classic()

clean_gilens %>%
  
  ggplot(aes(intgrp_align)) +
  
    # I use a histogram to display the underlying preference distribution.
  
    geom_histogram(aes(y = after_stat(count/sum(count))),
                   alpha = 0.5, 
                   bins = 12, 
                   position = "identity") +
  
    # I use the coefficients from my Model 4 output and the neutral points of 
    # the other two actors - 0.5 for average citizens and 0.5 got economic 
    # elites - to construct a line graph that shows how the predicted 
    # probability of policy change alters when moving from one point to another 
    # on the distribution of policy dispositions.
  
    geom_smooth(aes(y = -0.20999 + -0.86716*0.5 + 1.60551*0.5 + 0.48878*intgrp_align_norm)) +
    xlim(-24, 24) +
    
    # I customize the Y scales to have dual axes.
  
    scale_y_continuous(
      
      # I specify the name and limits of the lefthand axis.
      
      name = "Predicted probability of adoption",
      limits = c(0, 0.7),
      
      # I specify the name and limits of the second axis by transforming
      # the first axis.
      
      sec.axis = sec_axis(trans = ~.*0.5714, 
                          name = "Percent of cases (grey columns)")
    ) +
    labs(title = "Interest Group Alignments",
         x = "Net interest groups in support or opposition") +
    theme_classic()
```

```{r generate table 4}
# I run a multivariate model where the preferences of the median voter, the 
# preferences of the 90th percentile voter, and the interest group alignment
# cores of mass-based and business interest groups are predictors of policy 
# change.

model_5 <- lm(policy_change ~ pred50_norm + pred90_norm + mass_intgrp_align_norm + bus_intgrp_align_norm, data = clean_gilens)
summary(model_5)

# I use stargazer() to format a table of the regression output.

stargazer(model_5,
          type = "html", 
          title = "Table 4. The separate policy impact of business- oriented and mass-based interest groups",
          align = TRUE,
          dep.var.labels = c("Policy change"),
          covariate.labels = c("Average citizens’ preferences", 
                               "Economic elites’ preferences", 
                               "Mass-based interest groups",
                               "Business interest groups"),
          omit.stat=c("LL","ser","f", "adj.rsq", "n"), 
          no.space=TRUE,
          notes = c("*p<0.1; **p<0.05; ***p<0.01", "All predictors are scaled to range from 0 to 1. The dependent variable is the policy outcome, coded 1 if the proposed policy change took place within four years of the survey date and 0 if it did not. Predictors are the logits of the imputed percent of respondents at the fiftieth (“average citizens”) or ninetieth (“economic elites”) income percentile that favor the proposed policy change, and the Net Interest- Group Alignment Indices described in the text. Standard errors are asymptotically distribution-free, and all analyses reflect estimated measurement error in the predictors, as described in Appendix 2. N = 1,779."),
          notes.align = "l",
          notes.append = FALSE)
```

## Extension of the paper

Add a variable?

Logistic regression?

## Bibliography

Hlavac, Marek (2018). stargazer: Well-Formatted Regression and Summary Statistics
Tables. R package version 5.2.2. https://CRAN.R-project.org/package=stargazer


mass_intgrp_names <- 
  
if (names1 %in% mass_intgrp_names) {
  if
  table_2[mass_intgrp_stfav]
}

sum(names1 %in% names2)
mutate(mass_intgrp_stfav = sum(names1[names1 %in% names2] == 2))

mass_intgrp_names <- c("aarp", "american_bankers_association", "american_federationof_state_countyand_municipal_employees", "american_israel_public_affairs_committee",
            "american_legion", "christian_coalition", "international_brotherhoodof_teamsters", "national_rifle_association", 
            "national_rightto_life_committee", "united_auto_workersunion", "veteransof_foreign_warsofthe_u_s")
