---
title: "Replication Report"
author: "Monica Chang"
date: "4/24/2021"
output: 
  pdf_document:
    extra_dependencies: ["dcolumn"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(tidyverse)
library(janitor)
library(rstanarm)
library(stargazer)
library(haven)
library(modEvA)
library(dplyr)
library(statxp)
library(gt)
library(cowplot)
library(kableExtra)
```

## Overview of the original paper

### What question(s) did the author(s) ask?  

In the paper "Testing Theories of American Politics: Elites, Interest Groups, 
and Average Citizens" by Martin Gilens and Benjamin Page, the authors attempt to
answer the following question: Which of the four theoretical traditions in the 
study of American politics - Majoritarian Electoral Democracy, Economic-Elite 
Domination, Majoritarian Pluralism, Biased Pluralism - provides the most 
accurate picture of American politics?

These theories each offer different predictions over how much independent 
influence the median voter, economic elites, mass-based interest groups, and 
business-oriented interest groups have on public policy:

- Majoritarian Electoral Democracy posits that only the median voter has 
substantial independent influence on policy outcomes.

- Economic-Elite Domination posits that economic elites have the most
independent influence on policy outcomes, with the median voter and 
interest groups playing a smaller role.

- Majoritarian Pluralism posits that only organized interest groups will have
substantial independent influence on policy outcomes. In particular, this 
theory assumes that the stands of all interest groups, taken together, generally 
represent the preferences of average citizens. 

- Biased Pluralism posits that only organized interest groups will have
substantial independent influence on policy outcomes. In particular, this theory
predicts that business-oriented interest groups play a larger role.

Gilens and Page aim to test these four theoretical traditions within one
statistical model.

### How did they create their dataset? What were their variables of interest? 

To conduct this analysis, Gilens assembled a unique dataset of 1,779 instances 
between 1981 and 2002 in which a national survey of the general public asked a 
favor/oppose question about a proposed policy change.

For each survey question, Gilens' team did the following work to prepare a 
unique dataset:

- Use the available income breakdown for each survey question's 
respondents to estimate the opinions of respondents at the 10th income 
percentile, the 50th percentile, and the 90th percentile. 

- Record how many engaged powerful interest groups were strongly in favor of, 
somewhat in favor of, somewhat opposed, or strongly opposed to the proposed 
policy change in each question. This allowed Gilens to develop a measure of
Net Interest Group Alignment for all interest groups, mass-based interest 
groups, and business-oriented interest groups.

- Track whether the proposed policy change was actually adopted within 4 years
after the question was asked.

Predictor variables:

- Policy preferences at the 50th income percentile work as a proxy for the 
preferences of the median voter.

- Policy preferences at the 90th income percentile work as a proxy for the 
preferences of the economic elite.

- Net Interest Group Alignment scores work as a proxy for the preferences
of all interest groups, mass-based interest groups, and business-oriented 
interest groups.

Outcome variable:

- Whether or not the proposed policy change was adopted within 4 years.

### What analyses did they run? What models did they use?

First, they create a correlation matrix (Table 2) to assess the correlation
between independent variables. This helps test one of Majoritarian Pluralism's
predictions - that the stands of all interest groups are positively correlated 
with the preferences of average citizens. 

Then, they use structural equation modeling to regress policy outcome on
the preferences of the median voter, the preferences of the economic elite,
and the preferences of all interest groups (Table 3) to evaluate support 
for Majoritarian Electoral Democracy, Economic Elite Domination, and interest
group pluralism.

Finally, they regress policy outcome on the preferences of the median voter, 
the preferences of the economic elite, the preferences of mass-based interest 
groups, and the preferences of business-oriented interest groups to distinguish
between Majoritarian Pluralism and Biased Pluralism (Table 4).

### Why did they choose structural equation modeling?

They chose structural equation modeling to correct for correlated measurement 
error. Because the measures for two independent variables - the preferences of 
the 50th percentile voter and the preferences of the 90th percentile voter - 
come from a single data source (the same set of national opinion surveys), it 
is highly likely that the errors of these two measures are positively correlated 
with each other. These correlated errors tend to produce upwardly biased
statistical relationships between the two independent variables. This decision 
is further explained in Appendix 2 of their paper (see the documents folder
of my Github repository to read this appendix in detail).

### What did they conclude?

Gilens and Page conclude that economic elites and business-oriented interest 
groups have substantial independent impacts on U.S. policy, while mass-based 
interest groups and average citizens have little or no independent influence. 
Their results support theories of Economic-Elite Domination and Biased 
Pluralism.

## Overview of my replication

Note: Because structural equation modeling is outside the scope of this course, 
I use ordinary least squares regression instead. This difference in model choice 
means that my tables and figures are all slightly different from those shown in 
the paper. Besides this modeling difference, I fully replicate the four tables 
and one figure that appear in the paper.

### Data cleaning

For my replication, I first cleaned Gilens' publicly provided data according 
to the criteria laid out in the paper (excluding questions without an
income breakdown, where the policy was partially adopted, or if the policy 
change required a Constitutional amendment or Supreme Court ruling).

After filtering the dataset to only include the 1,779 survey questions that
met the criteria, I cleaned the dataset to prepare it for modeling:

- First, I turned the outcome variable into a dichotomous variable so that it was coded 1 
if the proposed policy change took place within four years of the survey date 
and 0 if it did not.

- I took the logits of the imputed percent of respondents at the 
50th and 90th income percentiles that favored the proposed policy change. 

- Then, I calculated Net Interest Group Alignment scores for all interest groups,
mass-based interest groups, and business-oriented interest groups for each
survey question by using the formula described in the paper:

Net Interest-Group Alignment = ln(# Strongly Favor + [0.5 * # Somewhat Favor] + 1) - ln(# Strongly Oppose + [0.5 * # Somewhat Oppose] + 1)

- Finally, I standardized all predictors to be on a scale of 0 to 1. 

\newpage

### Table 1 Replication

This table summarizes the main hypotheses of interest based on the four ideal 
types of theory. 

```{r generate table 1}
# I initialize the appropriate values for sets of actors that would 
# have independent influence for each of the theories of U.S. politics.

table_1 <- tibble(theories = c("Majoritarian Electoral Democracy", 
                               "Dominance by Economic Elites", 
                               "Majoritarian Pluralism", 
                               "Biased Pluralism"),
                  pred50_norm = c("Y", "y", "y", "n"),
                  pred90_norm = c("n", "Y", "n", "n"),
                  intgrp_align_norm = c("n", "y", "Y", "y"),
                  mass_intgrp_align_norm = c("n", "n", "Y", "y"),
                  bus_intgrp_align_norm = c("n", "y", "Y", "Y"))

# I format the table using functions from the kableExtra package.

kbl(table_1, 
    caption = "Theoretical predictions concerning the independent influence of sets of actors upon policy outcomes", 
    col.names = c("Theory (ideal type)",
                  "Average Citizens",
                  "Economic Elites",
                  "All Interest Groups",
                  "Mass Interest Groups",
                  "Business Interest Groups"),
    booktabs = T) %>%
  kable_styling(latex_options = c("scale_down")) %>%
  add_header_above(c(" " = 1, "Sets of Actors" = 5)) %>%
  footnote(general = "n = little or no independent influence \n y = some independent influence \n Y = substantial independent influence",
           footnote_as_chunk = T)

```

For example, for the theory of Majoritarian Electoral Democracy to be supported,
we would expect to see only the average citizen have substantial independent 
influence on the policy outcome.

For the theory of Dominance by Economic Elites to be supported, we would expect 
to see economic elites have the most substantial independent influence on the
policy outcome. We would also expect some independent influence by average 
citizens and interest groups - especially business-oriented interest groups.

The rest of the rows can be interpreted in a similar manner.

\newpage

### Table 2 Replication

This table shows the correlations between independent variables. 

```{r generate table 2}
clean_gilens <- read_csv(file = "clean_data/clean_gilens.csv")

# I select the five predictors relevant for Table 2.

table_2 <- clean_gilens %>%
  dplyr::select(pred50_norm, pred90_norm, intgrp_align_norm, 
                mass_intgrp_align_norm, bus_intgrp_align_norm)

# I use the corstars() function to compute a correlation matrix with stars to
# indicate significant correlations.

table_2_correlation <- corstars(table_2) %>%
  mutate(predictors = c("Average citizens’ preferences", "Economic elites’ preferences", 
                        "All interest groups", "Mass public interest groups", 
                        "Business interest groups")) %>%
  mutate(bus_intgrp_align_norm = "") %>%
  select(predictors, pred50_norm:mass_intgrp_align_norm, bus_intgrp_align_norm)

rownames(table_2_correlation) <- NULL

# I format the table using functions from the kableExtra package.

kbl(table_2_correlation, 
    caption = "Correlations among independent variables", 
    col.names = c(" ",
                  "Average citizens’ preferences",
                  "Economic elites’ preferences",
                  "All interest groups",
                  "Mass public interest groups",
                  "Business interest groups"),
    booktabs = T) %>%
    kable_styling(latex_options = c("scale_down")) %>%
    footnote(general = "***p<.001; n = 1779.")
```

The authors get the following findings from Table 2 in the original paper:

- The preferences of average citizens are positively and highly 
correlated with the preferences of economic elites. This means
average citizens might often be observed to get their preferred policy outcomes
even if they had no independent effect on policy making. 

- Net interest-group stands are not substantially correlated with the 
preferences of average citizens. Even mass-based interest groups have a very
modest correlation with the preferences of average citizens. This does *not*
support Majoritarian Pluralism's prediction - that organized interests 
taken together are good representatives of citizens as a whole.

- Business-oriented interest groups have a negative correlation with the 
preferences of average citizens. This suggests that business-oriented interest
groups are opposed to the interests of average citizens.

- There is no association between the preferences of economic elites and 
the alignments of either mass-based or business-oriented groups. While the 
latter finding may be surprising, this may be because economic elites broadly 
prefer lower government spending whereas interest groups generally lobby for 
increased spending for their specific industries.

The correlation coefficients in my table are slightly different from those in
the paper because I do not correct for correlated measurement error using
SEM modeling (SEM modeling is outside the scope of this course). However, my 
findings are generally quite similar except for the following differences:

- The correlation coefficient between the preferences of average citizens and the 
preferences of economic elites is much higher (0.94) than what Gilens and Page
get (0.78). This high correlation could be partly due to factors like sampling, 
question wording, question order, topics in the news at the time the survey 
was in the field, and so on.

- The correlation coefficient between the preferences of all interest groups and 
the preferences of mass-based interest groups is 0.19, whereas Gilens and Page
get 0.47. 

- The correlation coefficient between the preferences of all interest groups and 
the preferences of business-oriented interest groups is 0.64, whereas Gilens and 
Page get 0.96.

- The correlation coefficient between the preferences of mass-based interest 
groups and the preferences of business-oriented interest groups is -0.21 and 
statistically significant, whereas Gilens and Page get a coefficient of -0.05 
that is not statistically significant.

\newpage

### Table 3 Replication

I replicate Table 3. This table reports bivariate results for Models 1-3. Each of 
three independent variables (median voter, economic elite, all interest groups) 
is modeled separately as the sole predictor of policy change. The table reports
multivariate results for Model 4. All three independent variables are included 
in the multivariate model. I include the standardized coefficients for Model 4
in the footnote.

Note: Because structural equation modeling is outside the scope of this course, 
I use ordinary least squares regression instead. I am able to check whether my 
results are correct based on Table A1 in Appendix 2 - an OLS analysis provided 
by Gilens that is parallel to the structural equation model presented in 
Table 3. 

```{r generate table 3}
# I run a model where the preferences of the median voter is the sole predictor 
# of policy change.

model_1 <- lm(policy_change ~ pred50_norm, data = clean_gilens)

# I run a model where the preferences of the 90th percentile voter is the sole 
# predictor of policy change.

model_2 <- lm(policy_change ~ pred90_norm, data = clean_gilens)

# I run a model where the interest group alignment is the sole predictor of 
# policy change.

model_3 <- lm(policy_change ~ intgrp_align_norm, data = clean_gilens)

# I run a multivariate model where the preferences of the median voter, the 
# preferences of the 90th percentile voter, and the interest group alignment
# are predictors of policy change.

model_4 <- lm(policy_change ~ pred50_norm + pred90_norm + intgrp_align_norm, data = clean_gilens)

# I use the procedure on p. 56 in the textbook to standardize my coefficients.

clean_gilens_2 <- clean_gilens %>%
  
  # I subtract the mean and divide by 2 standard deviations.
  
  mutate(z.pred50 = (pred50_norm - mean(pred50_norm))/(2*sd(pred50_norm)),
         z.pred90 = (pred90_norm - mean(pred90_norm))/(2*sd(pred90_norm)),
         z.intgrp = (intgrp_align_norm - mean(intgrp_align_norm))/(2*sd(intgrp_align_norm)))

model_5 <- lm(policy_change ~ z.pred50 + z.pred90 + z.intgrp, data = clean_gilens_2)

```

```{r, results = "asis"}
# I use stargazer() to format a table with 3 regressions side-by-side.

stargazer(model_1, model_2, model_3, model_4, 
          title = "Policy outcomes and the policy preferences of average citizens, economic elites, and interest groups",
          align = TRUE,
          dep.var.labels = c("Policy change"),
          column.labels = c("Model 1", "Model 2", "Model 3", "Model 4"),
          covariate.labels = c("Preferences of average citizens", 
                               "Preferences of economic elites", 
                               "Alignment of interest groups"),
          omit.stat=c("LL","ser","f", "adj.rsq", "n"), 
          no.space=TRUE,
          notes = c("*p<0.1; **p<0.05; ***p<0.01", 
                    "All predictors are scaled to range from 0 to 1. The",
                    "dependent variable is the policy outcome, coded 1 if the", 
                    "proposed policy change took place within four years of the",
                    "survey date and 0 if it did not. Predictors are the logits",
                    "of the imputed percent of respondents at the fiftieth",
                    "(“average citizens”) or ninetieth (“economic elites”)",
                    "income percentile that favor the proposed policy change", 
                    "and the Net Interest-Group Alignment Index described in",
                    "the text. The standardized coefficients for model 4 in",
                    "this table are -0.22, .41, and .15 for average citizens",
                    "economic elites, and interest groups, respectively.",
                    "N = 1,779."),
          notes.align = "l",
          notes.append = FALSE)
```

The authors get the following findings from Table 3 in the original paper:

- When taken separately, each independent variable is strongly, positively, and 
significantly related to policy change.

- In the multivariate model, the independent influence of the median voter's 
preferences is not significant and near-zero. This rejects the theory of 
Majoritarian Electoral Democracy.

- Economic elites have a substantial, highly significant, independent influence
on policy outcomes.

- Organized interest groups are found to have substantial independent influence 
on policy outcomes.

Based on my replicated Table 3, my coefficient values for the bivariate 
regression models (Models 1-3) are quite similar to those in the paper. However,
my results are significantly different when it comes to the multivariate model
(Model 4):

- Whereas the regression coefficient representing the median voter's influence 
in the original paper is statistically and substantively insignificant (0.03), 
the coefficient representing the median voter's influence is negative, 
substantial, and statistically significant in my replication (-0.87). In other 
words, my replication indicates that a median voter's preference in 
favor of a policy change is associated with a *negative* influence on the policy
outcome. It would seem that Gilens and Page's correction for correlated 
measurement error is crucial to get to their key finding - the fact the median 
voter has no significant independent influence.

- The regression coefficient representing the economic elite's influence is much
higher in my replication (1.61) than in the original paper (0.76).

- The regression coefficient representing interest group influence is slightly 
lower in my replication (0.49) than in the original paper (0.56).

\newpage

### Figure 1 Replication

Because the magnitudes of the coefficients reported in Table 3 are hard to 
interpret, the authors visualize the relative influence of each set of actors. 
They compare how the predicted probability of policy change differs when moving 
from one point to another on their distributions of policy dispositions, while 
holding other actors’ preferences constant at their neutral points (50% 
favorable for average citizens and for economic elites, and a net interest-group 
alignment score of 0). These changing probabilities, based on the coefficients 
in Model 4 of Table 3, are line-graphed in Figure 1 along with bar graphs of 
the underlying preference distributions.

```{r figure 1}
# I pull out the coefficients from Model 4 to use in the plots.

b <- coef(model_4)

# I create the first plot in Figure 1.

plot_1 <- clean_gilens %>%
  
  # I multiply pred50_sw to obtain a number that represents the percentage 
  # of median voters in favor of a proposed policy change.
  
  mutate(pred50_sw_percent = 100*pred50_sw) %>%
  ggplot(aes(pred50_sw_percent)) +
  
    # I use a histogram to display the underlying preference distribution.
  
    geom_histogram(aes(y = after_stat(count/sum(count))),
                   alpha = 0.5, 
                   bins = 10, 
                   position = "identity") +
  
    # I use the coefficients from my Model 4 output and the neutral points of 
    # the other two actors - 0.5 for economic elites and 0 for interest group
    # alignment - to construct a line graph that shows how the predicted 
    # probability of policy change alters when moving from one point to another 
    # on the distribution of policy dispositions.
  
    geom_line(aes(y = invlogit(b[1] + b[2]*pred50_norm + b[3]*0.5 + b[4]*0))) +
    xlim(0, 100) +
    
    # I customize the Y scales to have dual axes.
  
    scale_y_continuous(
      
      # I specify the name and limits of the lefthand axis.
      
      name = "Predicted \n probability \n of adoption",
      limits = c(0, 0.7),
      
      # I specify the name and limits of the second axis by transforming
      # the first axis.
      
      sec.axis = sec_axis(trans = ~.*0.5714, 
                          name = "Percent of cases \n (grey columns)")
    ) +
    labs(title = "Average Citizens' Preferences",
         x = "Percent favoring proposed policy change") +
    theme_classic() +
    theme(axis.text=element_text(size=6))

# I create the second plot in Figure 1.

plot_2 <- clean_gilens %>%
  
  # I multiply pred90_sw to obtain a number that represents the percentage 
  # of median voters in favor of a proposed policy change.
  
  mutate(pred90_sw_percent = 100*pred90_sw) %>%
  ggplot(aes(pred90_sw_percent)) +
  
    # I use a histogram to display the underlying preference distribution.
  
    geom_histogram(aes(y = after_stat(count/sum(count))),
                   alpha = 0.5, 
                   bins = 10, 
                   position = "identity") +
  
    # I use the coefficients from my Model 4 output and the neutral points of 
    # the other two actors - 0.5 for average citizens and 0 for interest group
    # alignment - to construct a line graph that shows how the predicted 
    # probability of policy change alters when moving from one point to another 
    # on the distribution of policy dispositions.
  
    geom_line(aes(y = invlogit(b[1] + b[2]*0.5 + b[3]*pred90_norm + b[4]*0))) +
    xlim(0, 100) +
    
    # I customize the Y scales to have dual axes.
  
    scale_y_continuous(
      
      # I specify the name and limits of the lefthand axis.
      
      name = "Predicted \n probability \n of adoption",
      limits = c(0, 0.7),
      
      # I specify the name and limits of the second axis by transforming
      # the first axis.
      
      sec.axis = sec_axis(trans = ~.*0.5714, 
                          name = "Percent of cases \n (grey columns)")
    ) +
    labs(title = "Economic Elites' Preferences",
         x = "Percent favoring proposed policy change") +
    theme_classic() +
    theme(axis.text=element_text(size=6))

# I create the third graph in Figure 1.

plot_3 <- clean_gilens %>%
  
  ggplot(aes(intgrp_align)) +
  
    # I use a histogram to display the underlying preference distribution.
  
    geom_histogram(aes(y = after_stat(count/sum(count))),
                   alpha = 0.5, 
                   bins = 12, 
                   position = "identity") +
  
    # I use the coefficients from my Model 4 output and the neutral points of 
    # the other two actors - 0.5 for average citizens and 0.5 got economic 
    # elites - to construct a line graph that shows how the predicted 
    # probability of policy change alters when moving from one point to another 
    # on the distribution of policy dispositions.
  
    geom_line(aes(y = invlogit(b[1] + b[2]*0.5 + b[3]*0.5 + b[4]*intgrp_align_norm))) +
    xlim(-4, 4) +
    
    # I customize the Y scales to have dual axes.
  
    scale_y_continuous(
      
      # I specify the name and limits of the lefthand axis.
      
      name = "Predicted \n probability \n of adoption",
      limits = c(0, 0.7),
      
      # I specify the name and limits of the second axis by transforming
      # the first axis.
      
      sec.axis = sec_axis(trans = ~.*0.5714, 
                          name = "Percent of cases \n (grey columns)")
    ) +
    labs(title = "Interest Group Alignments",
         x = "Net interest groups in support or opposition") +
    theme_classic() +
    theme(axis.text=element_text(size=6))

# I stack all 3 plots on top of each other to form Figure 1.

title <- ggdraw() + draw_label("Figure 1. Predicted probability of policy adoption (dark lines, left axes) \n by policy disposition; the distribution of preferences (gray columns, right axes)", fontface='bold', size = 12)

plot_grid(
  title, plot_1, plot_2, plot_3,
  ncol = 1,
  nrow = 4,
  rel_heights = c(0.4, 1, 1, 1)
)
```

The authors get the following findings from Figure 1 in the original paper:

- When one holds the preferences of interest groups and economic elites constant, 
it makes very little difference what the average citizen prefers. The 
probability of policy change is around the same (around 0.3) whether a tiny
minority or a large majority of average citizens favor a proposed policy change.

- When one holds the preferences of interest groups and average citizens constant, 
it makes a large difference what economic elites prefer. When 1 out of 5 
economic elites favor a policy, it is adopted 18% of the time. When 4 out of 5
economic elites favor a policy, it is adopted 45% of the time.

- When one holds the preferences of average citizens and economic elites constant, 
it makes a large difference what interest groups prefer. When the Net Interest
Group Alignment score is -2, it is adopted 21% of the time. When the Net Interest
Group Alignment score is +2, it is adopted 39% of the time.

Based on my replicated Figure 1, I get slightly different results. This makes
sense considering the fact that my coefficients for Model 4 of Table 3 
are significantly different from the original paper. I find that:

- When one holds the preferences of interest groups and economic elites constant, 
it makes a large difference what the average citizen prefers. When 20% of 
average citizens favor a proposed policy change, it is adopted 60% of the time.
When 80% of average citizens favor a proposed policy change, it is adopted 47% 
of the time.

- When one holds the preferences of interest groups and average citizens constant, 
it makes a large difference what economic elites prefer. When 1 out of 5 
economic elites favor a policy, it is adopted 42% of the time. When 4 out of 5
economic elites favor a policy, it is adopted 65% of the time.

- When one holds the preferences of average citizens and economic elites constant, 
it makes a large difference what interest groups prefer. When the Net Interest
Group Alignment score is -2, it is adopted 31% of the time. When the Net Interest
Group Alignment score is +2, it is adopted 76% of the time.

\newpage

### Table 4 Replication

This table reports results for a multivariate analysis regressing the policy
outcome on the preferences of average citizens, the preferences of economic 
elites, the preferenes of mass-based interest groups, and the preferences of
business-oriented interest groups. 

```{r generate table 4}
# I run a multivariate model where the preferences of the median voter, the 
# preferences of the 90th percentile voter, and the interest group alignment
# cores of mass-based and business interest groups are predictors of policy 
# change.

model_6 <- lm(policy_change ~ pred50_norm + pred90_norm + mass_intgrp_align_norm + bus_intgrp_align_norm, data = clean_gilens)
```

```{r, results = "asis"}
# I use stargazer() to format a table of the regression output.

stargazer(model_6,
          title = "The separate policy impact of business- oriented and mass-based interest groups",
          align = TRUE,
          dep.var.labels = c("Policy change"),
          covariate.labels = c("Average citizens’ preferences", 
                               "Economic elites’ preferences", 
                               "Mass-based interest groups",
                               "Business interest groups"),
          omit.stat=c("LL","ser","f", "adj.rsq", "n"), 
          no.space=TRUE,
          notes = c("*p<0.1; **p<0.05; ***p<0.01", 
                    "All predictors are scaled to range from 0 to 1. The",
                    "dependent variable is the policy outcome, coded 1 if the", 
                    "proposed policy change took place within four years of the",
                    "survey date and 0 if it did not. Predictors are the logits",
                    "of the imputed percent of respondents at the fiftieth",
                    "(“average citizens”) or ninetieth (“economic elites”)",
                    "income percentile that favor the proposed policy change", 
                    "and the Net Interest-Group Alignment Index described in",
                    "the text.",
                    "N = 1,779."),
          notes.align = "l",
          notes.append = FALSE)
```

The authors get the following findings from Table 4 in the original paper:

- The coefficients for both mass-based and business-oriented interest groups
are positive and highly significant - but the coefficient for business groups 
is nearly twice as large as that for the mass groups. This supports the 
predictions of Biased Pluralism theories. Gilens and Page suggest that the 
advantage of business-oriented interest groups comes down to two factors: 
numerical advantage and relative cohesion. There are around twice as many 
business-oriented groups engaged on each policy issue. In addition, business 
groups are rarely found  on both sides of a proposed policy change.

- The findings in Table 3 are confirmed in this analysis. The preferences of 
average citizens are still near-zero and not significant. Economic elites 
still have a substantial, positive, independent impact.

My coefficients for both mass-based and business-oriented interest groups are
fairly similar to the original paper. However, because I use OLS regression 
instead of SEM modeling, I get slightly different results for the other
two coefficients. My results here are consistent with my observations
for my replication's Table 3.

- My coefficient representing the influence of average citizens was substantial,
negative, and significant (-0.895) whereas the same coefficient in the original
paper was positive, near-zero, and not significant (0.05). 

- My coefficient representing the influence of economic elites was (1.662) 
was much higher than the same coefficient in the original paper (0.78). 

\newpage

## Extension of the paper

The original paper does not consider the impact of party control on policy 
outcomes. Intuitively, it would make sense for party control to have some 
influence over governmental responsiveness to voters of different income
percentiles. In general, the Democratic Party is associated with policies that
are more beneficial for low-income individuals and the Republican Party is
associated with policies that are more beneficial for high-income individuals
and corporations. 

This is an example of omitted variable bias. When a statistical model leaves
out one or more relevant variables, the model may attribute the effect of the 
missing variables to those that were included.

I extended the replication by adding in a variable to indicate whether
the party of the President in the year of the survey question was a Democrat or 
a Republican. I was curious about whether Democratic or Republican control of 
the presidency was more associated with policy change. 

```{r extension part 1}

clean_gilens_extension <- clean_gilens %>%
  
  # I add a variable to indicate the party of the President in any given year.
  # 1 is Republican, 0 is Democrat.
  
  mutate(party_pres = if_else(year >= 1981 & year <= 1992 | 
                              year >= 2001, 1, 0))

# I run a multivariate model where the preferences of the median voter, the 
# preferences of the 90th percentile voter, net interest group alignment,
# and party of the President are predictors of policy change.

model_7 <- lm(policy_change ~ pred50_norm + pred90_norm + 
                intgrp_align_norm + party_pres, 
              data = clean_gilens_extension)

# I use invlogit() to convert to predicted probabilities for policy change to
# be adopted. This makes the coefficient more interpretable.

c = coef(model_7)
# invlogit(c[1] + c[2]*0.5 + c[3]*0.5 + c[4]*0 + c[5]*0)
# invlogit(c[1] + c[2]*0.5 + c[3]*0.5 + c[4]*0 + c[5]*1)
```

```{r, results = "asis"}
# I use stargazer() to format a table of the regression output.

stargazer(model_7,
          title = "The impact of the party of the President on policy outcomes",
          align = TRUE,
          dep.var.labels = c("Policy change"),
          covariate.labels = c("Average citizens’ preferences", 
                               "Economic elites’ preferences", 
                               "Alignment of interest groups",
                               "Republican president"),
          omit.stat=c("LL","ser","f", "adj.rsq", "n"), 
          no.space=TRUE,
          notes = c("*p<0.1; **p<0.05; ***p<0.01", 
                    "All predictors are scaled to range from 0 to 1. The",
                    "dependent variable is the policy outcome, coded 1 if the", 
                    "proposed policy change took place within four years of the",
                    "survey date and 0 if it did not. Predictors are the logits",
                    "of the imputed percent of respondents at the fiftieth",
                    "(“average citizens”) or ninetieth (“economic elites”)",
                    "income percentile that favor the proposed policy change", 
                    "and the Net Interest-Group Alignment Index described in",
                    "the text.",
                    "N = 1,779."),
          notes.align = "l",
          notes.append = FALSE)
```

Based on the regression output, the other coefficients are fairly similar to 
what they were in Table 3. To interpret the party_pres coefficient, we can 
convert to predicted probabilities - similar to what we did for Figure 1.

When one holds the preferences of the median voter, economic elites, and
interest groups constant, the party of the President is associated with
little predictive difference. When the president is Democratic, proposed policy
changes are adopted 53% of the time. When the president is Republican, proposed 
policy changes are adopted 55% of the time. The difference seems relatively 
small.

\newpage

I also extended the replication by adding in a statistical interaction between 
party of the President and preferences of the median voter. In *Unequal 
Democracy*, Larry Bartels finds that partisan control has a strong impact on 
the minimum wage (with the minimum wage keeping pace with inflation during 
Democratic administrations but declining during Republican administrations). 
This suggests that Democratic administrations are more responsive to the
median voter. I was curious about whether there was any differential 
responsiveness by the different parties to the median voter based on this 
dataset.

Of course, my analysis is fairly limited as I am not considering the party in
control of Congress, whether government is unified, and a whole host of other
potentially confounding factors.

```{r extension part 2}
# I run a multivariate model where the preferences of the median voter, the 
# preferences of the 90th percentile voter, net interest group alignment,
# party of the President, and an interaction between party of the President and
# preferences of average citizens are predictors of policy change.

model_8 <- lm(policy_change ~ pred50_norm + pred90_norm + intgrp_align_norm +
              party_pres + party_pres:pred50_norm, 
              data = clean_gilens_extension)

# I use invlogit() to convert to predicted probabilities for policy change to
# be adopted. This makes the coefficient more interpretable.

d = coef(model_8)
# invlogit(d[1] + d[2]*0.2 + d[3]*0.5 + d[4]*0 + d[5]*0 + d[6]*0)
# invlogit(d[1] + d[2]*0.8 + d[3]*0.5 + d[4]*0 + d[5]*0 + d[6]*0)

# invlogit(d[1] + d[2]*0.2 + d[3]*0.5 + d[4]*0 + d[5]*1 + d[6]*0.2)
# invlogit(d[1] + d[2]*0.8 + d[3]*0.5 + d[4]*0 + d[5]*1 + d[6]*0.8)
```

```{r, results = "asis"}

# I use stargazer() to format a table of the regression output.

stargazer(model_8,
          title = "The impact of the party of the President on policy outcomes",
          align = TRUE,
          dep.var.labels = c("Policy change"),
          covariate.labels = c("Average citizens’ preferences", 
                               "Economic elites’ preferences", 
                               "Alignment of interest groups",
                               "Republican president",
                               "Average citizens' preferences*Republican president"),
          omit.stat=c("LL","ser","f", "adj.rsq", "n"), 
          no.space=TRUE,
          notes =c("*p<0.1; **p<0.05; ***p<0.01", 
                    "All predictors are scaled to range from 0 to 1. The",
                    "dependent variable is the policy outcome, coded 1 if the", 
                    "proposed policy change took place within four years of the",
                    "survey date and 0 if it did not. Predictors are the logits",
                    "of the imputed percent of respondents at the fiftieth",
                    "(“average citizens”) or ninetieth (“economic elites”)",
                    "income percentile that favor the proposed policy change", 
                    "and the Net Interest- Group Alignment Index described in",
                    "the text.",
                    "N = 1,779."),
          notes.align = "l",
          notes.append = FALSE)
```

It appears that after adding this interaction, the coefficient for the 
party_pres coefficient decreases significantly and is no longer statistically 
significant (the standard error for the party_pres coefficient is larger in 
magnitude than the coefficient itself). 

However, the pred50_norm:party_pres coefficient is relatively statistically 
significant. This coefficient represents the change in pred50_norm
if the party of the President goes from Democrat to Republican. To more easily 
interpret this coefficient, we can convert to predicted probabilities.

When one holds the preferences of economic elites and interest groups constant, 
the party of the President is associated with some differential responsiveness 
to the median voter:

- When the president is Democratic and 1 out of 5 median voters favors 
the proposed policy change, the policy is adopted 60% of the time.
- When the president is Democratic and 4 out of 5 median voter favors 
the proposed policy change, the policy is adopted 45% of the time.
- When the president is Republican and 1 out of 5 median voters favors 
the proposed policy change, the policy is adopted 60% of the time.
- When the president is Republican and 4 out of 5 median voters favors 
the proposed policy change, the policy is adopted 50% of the time.

It appears that Republican presidents are more responsive to the median voter. 
This is surprising - as I would have expected a Democratic president
to be associated with more responsiveness to the median voter. In general,
Democratic administrations tend to align themselves with "the common man".

The implications of this are striking and present a new puzzle. Why is 
the Democratic Party, a party that frequently aligns itself with "the common 
man" and pushes for policies that are supposedly beneficial for lower-income
individuals, associated with *less* governmental responsiveness to the median 
voter? Some additional research questions inspired by this puzzle:

- In what policy domains were Republicans more responsive to the median voter?
- Is the Democratic Party out of touch with what the median voter wants in 
terms of social policies?
- Does the Democratic Party pass more downwardly redistributive policies?
- How different are preferences on redistributive policies between the median 
voter and affluent voters? 
- Does the average citizen prefer more downwardly redistributive policies?
- How many of the policies on the national agenda are redistributive?

## Conclusion

In this replication, I re-create all of the tables and figures of the original 
paper. The most important findings of the original paper were:

- Economic elites and business-oriented interest groups have substantial 
independent impacts on U.S. policy, while mass-based interest groups and 
the median voter have little or no independent influence. 
- Their results support theories of Economic-Elite Domination and Biased 
Pluralism.

I use OLS regression instead of SEM modeling (like Gilens and Page) because SEM 
modeling is outside the scope of this class. This poses a limitation for how
closely I am able to replicate their results. This difference in model choice 
results in one major difference from their original paper - the influence of 
the amedian voter. In my analyses, the preferences of the median voter had 
a substantial, *negative*, and significant influence on the policy outcome.
In other words, my replication indicates that a median voter's preference in 
favor of a policy change is associated with a *negative* influence on the 
proposed policy change getting adopted.

For my extension, I looked into whether Democratic or Republican control of 
the presidency was more associated with policy change. I found little influence.
I also looked into whether there was any differential responsiveness to the 
median voter by the different parties. I surprisingly found that
Republican presidents are more responsive to the median voter.

Future questions that could be investigated with this dataset:

- What is the impact of unified government on governmental responsiveness?
Are periods of unified government associated with greater likelihood that
proposed policy changes will get adopted?
- What would these results look like if this dataset was extended into the
present day? Given the impact of the filibuster, has governmental responsiveness
worsened?

## Data Source & Code Repository

The data for this paper can be found here: https://www.russellsage.org/datasets/economic-inequality-and-political-representation. 
All the code used to generate this report can be found here: https://github.com/monica-chang/gilens-replication.

## Bibliography

Bartels, Larry M. Unequal Democracy: The Political Economy of the New Gilded Age. New York; Princeton; Oxford: Princeton University Press, 2008. Accessed May 4, 2021. http://www.jstor.org/stable/j.ctt7t9ks.

Gilens, M., & Page, B. (2014). Testing Theories of American Politics: Elites, Interest Groups, and Average Citizens. Perspectives on Politics, 12(3), 564-581. doi:10.1017/S1537592714001595

O'Brien, Robert M. "Correcting Measures of Relationship between Aggregate-Level Variables." Sociological Methodology 21 (1994): 125-65

O'Brien, Robert M. "Correcting Measures of Relationship between Aggregate-Level Variables for Both Unreliability and Correlated Errors: An Empirical Example." Social Science Research 27, no. 2 (Jun 1998): 218-34.

